---
title: "01-data_cleaning-post-strat1"
author: "Jiaheng Li, Anni Lin, Yuechen Shen, Yuxin Yang"
date: "28/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Preamble ####
# Purpose: Prepare and clean the census data downloaded from *IPUMS USA, University of Minnesota, www.ipums.org.*
# Author: Jiaheng Li, Anni Lin, Yuechen Shen, Yuxin Yang
# Data: 28 October 2020
# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
# License: MIT
# Pre-requisites: 
# - Need to have downloaded the ACS data and saved it to inputs/data
# - Don't forget to gitignore it!

 
#### Workspace setup ####
```{r}
library(haven)
library(tidyverse)
library(dplyr)
# Read in the raw data.
setwd("~/Downloads/UofT 2018-2022/Academic/STA304/STA304_PS3")
raw_data <- read_dta("usa_00001.dta.gz")

# Add the labels
raw_data <- labelled::to_factor(raw_data)
```

# Keep some variables of interest
```{r}
reduced_census_data <- raw_data %>%
  dplyr::select(stateicp, 
                sex, 
                age, 
                race, 
                empstatd)
```

#### What's next? ####

## Here I am only splitting cells by age, but you 
## can use other variables to split by changing
## count(age) to count(age, sex, ....)
```{r}
reduced_census_data <- reduced_data %>%
  as.tibble() %>%
  dplyr::rename(state = stateicp, gender = sex, race_ethnicity = race, employment = empstatd) %>%
  filter(race_ethnicity != "two major races") %>%
  filter(race_ethnicity != "three or more major races") %>%
  filter(as.double(age) > 17) %>%
  filter(age != "less than 1 year old")

reduced_census_data$gender <- ifelse(reduced_census_data$gender=="male", "Male", "Female")
reduced_census_data$state <- str_to_title(reduced_census_data$state)

reduced_census_data$employment <- ifelse(reduced_census_data$employment=="at work", "Employed", 
                                ifelse(reduced_census_data$employment=="at work, public emerg", "Employed", 
                                ifelse(reduced_census_data$employment=="has job, not working", "Employed", 
                                ifelse(reduced_census_data$employment=="armed forces", "Employed", 
                                ifelse(reduced_census_data$employment=="armed forces--at work", "Employed", 
                                ifelse(reduced_census_data$employment=="armed forces--not at work but with job", "Employed", 
                                ifelse(reduced_census_data$employment=="unemployed", "Unemployed", 
                                ifelse(reduced_census_data$employment=="unemp, exper worker", "Unemployed", 
                                ifelse(reduced_census_data$employment=="unemp, new worker", "Unemployed", 
                                ifelse(reduced_census_data$employment=="not in labor force", "Not in labor force", 
                                ifelse(reduced_census_data$employment=="nilf, housework", "Not in labor force", 
                                ifelse(reduced_census_data$employment=="nilf, unable to work", "Not in labor force", 
                                ifelse(reduced_census_data$employment=="nilf, school", "Student", 
                                ifelse(reduced_census_data$employment=="nilf, other", "Not in labor force",  
                        NA  ))))))))))))))

#   group_by(state, gender) %>%
#   count()
```

```{r}
reduced_census_data <- reduced_census_data %>%
  group_by(state, gender, race_ethnicity) %>%
  count()

View(reduced_census_data)
```

```{r}
reduced_data <- 
  reduced_data %>%
  count(age) %>%
  group_by(age) 

reduced_data <- 
  reduced_data %>% 
  filter(age != "less than 1 year old") %>%
  filter(age != "90 (90+ in 1980 and 1990)")

reduced_data$age <- as.integer(reduced_data$age)

reduced_data
```


# Saving the census data as a csv file in my
# working directory
```{r}
write_csv(reduced_census_data, "/Users/yangyuxin/Downloads/UofT 2018-2022/Academic/STA304/STA304_PS3/census_data.csv")
```


         