---
title: "01-data_cleaning-survey1"
author: "Jiaheng Li, Anni Lin, Yuechen Shen, Yuxin Yang"
date: "28/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Preamble ####
# Purpose: Prepare and clean the survey data downloaded from voterstudygroup.org
# Author: Jiaheng Li, Anni Lin, Yuechen Shen, Yuxin Yang
# Data: 22 October 2020
# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
# License: MIT
# Pre-requisites: 
# - Need to have downloaded the data from X and save the folder that you're 
# interested in to inputs/data 
# - Don't forget to gitignore it!


#### Workspace setup ####
```{r}
library(haven)
library(tidyverse)
setwd("~/Downloads/UofT 2018-2022/Academic/STA304/STA304_PS3")
# Read in the raw data (You might need to change this if you use a different dataset)
raw_survey_data <- read_dta("ns20200625.dta")

# Add the labels
raw_survey_data <- labelled::to_factor(raw_survey_data)
```

```{r}
# Just keep some variables
reduced_survey_data <- 
  raw_survey_data %>% 
  dplyr::select(interest,
         registration,
         vote_2016,
         vote_intention,
         vote_2020,
         ideo5,
         employment,
         foreign_born,
         gender,
         census_region,
         hispanic,
         race_ethnicity,
         household_income,
         education,
         state,
         congress_district,
         age)
```

#### What else???? ####
# Maybe make some age-groups?
# Maybe check the values?
# Is vote a binary? If not, what are you going to do?
```{r}
reduced_survey_data<-reduced_survey_data %>%
  mutate(vote_trump = ifelse(vote_2020=="Donald Trump", 1, 0)) %>%
  #Convert state abbreviations to names
  mutate(state_name = state.name[match(state, state.abb)]) %>% 
  # select variables in interest
  dplyr::select(vote_2016, gender, age, race_ethnicity, employment, state_name, vote_trump) %>%
  na.omit()

reduced_survey_data$employment <- ifelse(reduced_survey_data$employment=="Full-time employed", "Employed", 
                                  ifelse(reduced_survey_data$employment=="Homemaker", "Not in labor force", 
                                  ifelse(reduced_survey_data$employment=="Retired", "Not in labor force", 
                                  ifelse(reduced_survey_data$employment=="Unemployed or temporarily on layoff", "Unemployed",
                                  ifelse(reduced_survey_data$employment=="Part-time employed", "Employed",
                                  ifelse(reduced_survey_data$employment=="Permanently disabled", "Not in labor force", 
                                  ifelse(reduced_survey_data$employment=="Student", "Student",
                                  ifelse(reduced_survey_data$employment=="Self-employed", "Employed", 
                                  ifelse(reduced_survey_data$employment=="Other:", "Other",
                        NA  )))))))))

reduced_survey_data$race_ethnicity <- ifelse(reduced_survey_data$race_ethnicity=="Asian (Asian Indian)", "Other Asian or Pacific Islander", 
                                      ifelse(reduced_survey_data$race_ethnicity=="Asian (Chinese)", "East Asian", 
                                      ifelse(reduced_survey_data$race_ethnicity=="Asian (Filipino)", "Other Asian or Pacific Islander", 
                                      ifelse(reduced_survey_data$race_ethnicity=="Asian (Japanese)", "East Asian",
                                      ifelse(reduced_survey_data$race_ethnicity=="Asian (Korean)", "East Asian",
                                      ifelse(reduced_survey_data$race_ethnicity=="Asian (Vietnamese)", "Other Asian or Pacific Islander", 
                                      ifelse(reduced_survey_data$race_ethnicity=="Asian (Other)", "Other Asian or Pacific Islander",
                                      ifelse(reduced_survey_data$race_ethnicity=="Pacific Islander (Native Hawaiian)", "Other Asian or Pacific Islander", 
                                      ifelse(reduced_survey_data$race_ethnicity=="Pacific Islander (Guamanian)", "Other Asian or Pacific Islander", 
                                      ifelse(reduced_survey_data$race_ethnicity=="Pacific Islander (Samoan)", "Other Asian or Pacific Islander", 
                                      ifelse(reduced_survey_data$race_ethnicity=="Pacific Islander (Other)", "Other Asian or Pacific Islander", 
                                      ifelse(reduced_survey_data$race_ethnicity=="Some other race", "Other",
                                      ifelse(reduced_survey_data$race_ethnicity=="White", "White", 
                                      ifelse(reduced_survey_data$race_ethnicity=="Black, or African American", "Black or African American", 
                                      ifelse(reduced_survey_data$race_ethnicity=="American Indian or Alaska Native", "American Indian or Alaska Native", 
                        NA  )))))))))))))))
```

```{r}
View(reduced_survey_data)
```


# Saving the survey/sample data as a csv file in my
# working directory
```{r}
write_csv(reduced_survey_data, "/Users/yangyuxin/Downloads/UofT 2018-2022/Academic/STA304/STA304_PS3/survey_data.csv")
```
