---
title: "Title of Your Report"
author: "Jiaheng Li, Anni Lin, Yuechen Shen, Yuxin Yang"
date: "27/10/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
library(visdat)

# Loading in the cleaned survey Data
survey_data <- read_csv("/Users/yangyuxin/Downloads/UofT 2018-2022/Academic/STA304/STA304_PS3/survey_data.csv")

# Loading in the cleaned census Data
census_data <- read_csv("/Users/yangyuxin/Downloads/UofT 2018-2022/Academic/STA304/STA304_PS3/census_data.csv") %>%
  na.omit()
```



# Title of your Report

## Name(s) of Author(s) 
## Date


# Model

Here we are interested in predicting the popular vote outcome of the 2020 American federal election (include citation). To do this we are employing a post-stratification technique. In the following sub-sections I will describe the model specifics and the post-stratification calculation.


## Model Specifics
We will be using a logistic regression model to model the proportion of voters who will vote for Donald Trump. This is a naive model. I will only be using age, which is recorded as a numeric variable, to model the probability of voting for Donald Trump. The simple linear regression model I am using is:

$$ y = \beta_0+\beta_1  x_{} + \epsilon$$

Where $y$ represents the proportion of voters who will vote for Donald Trump. Similarly, $\beta_0$ represents the intercept of the model, and is the probability of voting for Donald Trump at age 0. Additionally, $\beta_1$ represents the slope of the model. So, for everyone one unit increase in age, we expect a $\beta_1$ increase in the probability of voting for Donald Trump.
```{r}
# Creating the Model
model_logit_trump <- glm(vote_trump ~ gender + age + race_ethnicity + employment + state_name, data = survey_data, binomial)

# Model Results (to Report in Results section)
summary(model_logit_trump) #, or
# broom::tidy(model_logit_trump)
```

```{r}
census_data$logodds_estimate <- model_logit_trump%>%
  predict(newdata = census_data)

census_data$estimate <- exp(census_data$logodds_estimate)/(1+exp(census_data$logodds_estimate))

census_data <- census_data %>%
  mutate(elect_predict_prob = estimate * n) 
```

```{r}
Trump_prob <- sum(census_data$elect_predict_prob)/sum(census_data$n)
Trump_prob
```


### Joe Biden ###
```{r}
# Creating the Model
model_logit_biden <- glm(vote_biden ~ gender + age + race_ethnicity + employment + state_name, data = survey_data, binomial)

# Model Results (to Report in Results section)
summary(model_logit_biden) #, or
# broom::tidy(model_logit_biden)
```

```{r}
census_data$logodds_estimate2 <- model_logit_biden%>%
  predict(newdata = census_data)

census_data$estimate2 <- exp(census_data$logodds_estimate2)/(1+exp(census_data$logodds_estimate2))

census_data <- census_data %>%
  mutate(elect_predict_prob2 = estimate2 * n) 
```

```{r}
Biden_prob <- sum(census_data$elect_predict_prob2)/sum(census_data$n)
Biden_prob
```




